Highly Available Transactions: Virtues and Limitations

### 2013-04-23 writeup/hat-analysis.tex
bah
fixed
editing pass

HAT systems provide transactions with transactional availability and
either high availability or sticky high availability. They offer
substantial latency and availability benefits compared to traditional
distributed databases, yet they cannot achieve all the traditional
semantics. In this section, we delineate ACID, distributed
consistency, and session consistency levels which can be achieved with
high availability (transactional atomicity, variants of Repeatable
Read isolation, and many session guarantees), those with sticky high
availability (read your writes, PRAM and causal consistency) and the
properties which cannot be provided in a HAT system (those preventing
Lost Update and Write Skew, or with recency).  We present a full
summary of these results in Section~\ref{sec:hat-summary}.

@@

To begin, we present well-known semantics that can be achieved in HAT
systems. We offer proof-of-concept highly available algorithms but our
goal is \textit{not} to provide optimal or even efficient
implementations. These guarantees are fairly straightforward and have
been introduced---albeit with greater brevity and discussion
omitted---in a preliminary workshop paper~\cite{hat-hotos}. In our
examples, we exclusively consider read and write operations. We denote
a write of value $v$ to data item $d$ as $w_d(v)$ and a read from data
item $d$ returning $v$ as $r_d(v)$ and assume that all data items have
the null value, $\bot$, at database initialization. Unless otherwise
specified, all example transactions commit.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
PL-1,requiring that each transaction's writes are ordered consistently
according to a single total order on transactions. This prohibits
Adya's phenomenon $G0$, also called ``Dirty Writes''~\cite{adya}. If
two transactions write to the same set of data items, then the final
database state cannot contain the ``earlier'' transaction's writes to
the data items. For example, in the below example, $T_3$ should
\emph{eventually} only read $a=b=1$ or $a=b=2$ but not $a=2, b=1$ or
$a=1, b=2$:
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\\[-2em]
\end{align*}
All later properties will strengthen PL-1.

@@

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs. Implementations differ,
with some based on long-duration exclusive locks and short-duration
read locks~\cite{gray-isolation} and others based on multiple
versions. The implementations typically have recency and monotonicity
properties beyond the simple meaning of the name, which is what is
expressed in the implementation-agnostic definition: under Read
Committed, transactions should do not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's PL-2, and formalized by prohibiting Adya's $G1a-c$
(or ANSI's $P1$, or ``broad'' $P1$ (2.2) from Berenson et al.). For
instance, in the example below, $T_3$ should never see $a=1$, and, if
$T_2$ aborts, $T_3$ should never read $a=3$:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\\[-2em]
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
transaction never writes uncommitted data to the database, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. This implementation does not provide recency or
monotonicity guarantees but satisfies the implementation-agnostic
definition.

``Repeatable Read'' isolation is a confusing property. Some IBM
products use the term for fully serializable
isolation~\cite{hat-hotos}. Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all interpret
Repeatable Read as providing serializability for all operations except
for predicate-based reads (and so Repeatable Read is identical to
serializability in a key-value store without predicate-based
access). In this model, the \textit{Phantom Problem}, whereby two
successive predicate-based reads return different
data~\cite{gray-isolation}, is the only kind of non-serializable
behavior allowed. When referring to ``Repeatable Read,'' we will
follow these definitions, which, as we will soon show, are for a
property that can be offered in a highly available way.

@@

A stronger achievable property asks that a consistent cut spans both a
transaction's predicate reads (e.g., \texttt{SELECT WHERE}) as well as
its item reads.  We call this \textit{Predicate Cut
  Isolation}\footnote{Oracle provides an isolation level called
  ``Statement Level Read Consistency''~\cite{adya}; this is analogous
  to Predicate Cut Isolation at the level of a single operation within
  a multi-operation transaction. We do not consider this isolation
  model here as it is subsumed by standard Snapshot Isolation and we
  believe our discussion is easily extended to incorporate it.} and it
prohibits Phantoms. Predicate Cut Isolation is also achievable in HAT
systems via similar multi-versioning as above.

@@

Perplexingly, discussions of TA are absent from existing ones of
weak isolation. This is perhaps again due to the single-node
context in which prior work was developed: on a single server (or a
fully replicated database), TA is achievable via lightweight locking
and/or local concurrency control over data items~\cite{gstore}. In
contrast, in a distributed environment, TA over arbitrary groups of
non-colocated items is considerably more difficult to achieve with
high availability: servers need to know when all of a transaction's
updates are present on their respective replicas. However, replicas do
not need to agree on when to reveal a new value to clients, which
would require consensus; accordingly, TA only requires reliable
broadcast (with some additional client metadata) and is
achievable in HAT systems. We omit a full discussion of the algorithm
and instead refer the reader to the extended version of this
paper~\cite{hat-tr}.  
%space{.5em}

@@

Our models have not yet considered interactions \textit{across}
transactions: we have not guaranteed any ordering or continuity
between transactions, other than that there exists some arbitrary
ordering (via Read Uncommitted). In the distributed systems
literature, many useful \textit{safety} guarantees span multiple
application actions. In particular, \textit{session guarantees} are
used to describe guarantees across transactions within a given
\textit{session}, ``an abstraction for the sequence of...operations
performed during the execution of an
application''~\cite{sessionguarantees}. Informally a session describes
a context that should persist between transactions: for example, on a
social networking site, all of a user's transactions submitted between
``log in'' and ``log out'' operations might form a session.

@@

\vspace{.5em}\noindent\textit{{Monotonic reads}} requires that, within
a session, subsequent reads to a given object ``never return any
previous values''; reads from each item progress forward in a total
order. The ordering of reads should respect any externally observable
total ordering on transactions.

\vspace{.5em}\noindent\textit{{Monotonic writes}} requires that each
session's writes be applied to any replica in the order they were
submitted by the client. Any order on transactions should also be
consistent with any precedence that a global observer would see.

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions
in succession:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: r_x(a)
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client executes $T_2$ against the same (partitioned)
server, then it will be able to read its writes. However, if the
network topology shifts and the client can only contact a different
server that is partitioned from the server that executed $T_1$, then
the client will be unable to read its own writes and the system will
have to either stall indefinitely to allow the client to read her
writes (violating transactional availability) or will have to
sacrifice read your writes guarantees. However, if the client remains
sticky with the server that executed $T_1$, then we can disallow this
scenario. Accordingly, read your writes, and, by proxy, causal
consistency and PRAM require stickiness. Read your writes is ``free''
in a sticky system, while the remaining causality and PRAM guarantees
can be accomplished with the algorithms for achieving the remaining
session guarantees.

@@

\vspace{0.5em}
\noindent{\textbf{Consistency}} A HAT system can make limited
consistency guarantees. It can often execute commutative and logically
monotonic~\cite{calm} operations without the risk of invalidating
 application-level integrity constraints. Our goal in
this paper is not to sketch the entire space of consistency models
that are achievable (see Section~\ref{sec:futurework}). We
specifically evaluate TPC-C transaction semantics under HAT
consistency guarantees in Section~\ref{sec:evaluation}.

\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
transactions' effects survive $F$ server faults requires at least
$F+1$ correct nodes.

@@

In the words of Berenson et al., \textit{Lost Update} occurs when one
transaction $T1$ reads a given data item, a second transaction $T2$
updates the same data item, then $T1$ modifies the data item based on
its original read of the data item, ``missing'' or ``losing'' $T2$'s
newer update. Consider a database containing only the following
transactions:
\begin{align*}
\small
T_1 &: r_x(a) w_x(a+2)
\\T_2 &: w_x(2)
\end{align*}
If $T_1$ reads $a=1$ but $T_2$'s write to $x$ precedes $T_1$'s write
operation, then the database will end up with $a=3$, a state that
could not have resulted in a serial execution due to from $T_2$'s
``Lost Update.''

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:evaluation}) but, as we show in the extended version of
  the paper~\cite{hat-tr}, the problem persists in the general case.}
To prevent this from happening, either $T_1$ or $T_2$ should not have
committed. Each client's respective server might try to detect that
another write occurred, but this requires knowing the version of the
latest write to $x$. In our example, this reduces to a requirement for
linearizability, which is, via Gilbert and Lynch's proof of the CAP
Theorem, provably unachievable with high
availability~\cite{gilbert-cap}.

@@

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation, and Cursor Stability guarantees are all unavailable.
Repeatable Read (in the sense of Gray~\cite{gray-isolation}, Berenson
et al.~\cite{ansicritique}, or Adya~\cite{adya}) and One-Copy
Serializability need to prevent both Lost Update and Write Skew; this
means that they are also inherently unavailable.

@@

Data storage systems make various recency guarantees on reads of data
items. As we have discussed, one of the most famous is
linearizability~\cite{herlihy-art}, which states that reads will
return the last completed write to a data item, and there are several
other (weaker) variants such as safe and regular register
semantics. When applied to transactional semantics, the combination of
one-copy serializability and linearizability is called \textit{strong
  (or strict) one-copy serializability}~\cite{adya} (e.g., Google's
Spanner~\cite{spanner}). It is also common, particularly in systems
that allow reading from masters and slaves, to provide a guarantee
such as ``read a version that is no more than five seconds out of
date'' or similar. Unfortunately, an indefinitely long partition can
force an available system to violate any recency bound, so recency
bounds are not enforceable in HAT systems.

@@

We summarize our results in Table~\ref{table:hatcompared}. A wide
range of isolation levels are achievable in a HAT systems, including
transactional atomicity, cut isolation, and several session
guarantees. With sticky availability, a system can achieve read your
writes guarantees and PRAM and causal consistency. However, many other
prominent models, such as Snapshot Isolation, One-Copy
Serializability, and Strict Serializability cannot be achieved due to
the inability to prevent Lost Update and Write Skew phenomena.

@@

To the best of our knowledge, this is the first unification of
database ACID, distributed consistency, and session guarantee
models. Interestingly, strong one-copy serializability subsumes all
other models, while considering the (large) power set of all
compatible models (e.g., the diagram depicts 96 possible highly
available combinations) hints at the vast expanse of consistency
models found in the literature. This taxonomy  is \textit{not}
exhaustive, but we believe it lends substantial clarity into the
relationships between a large subset of the prominent ACID and
distributed consistency models. Additional read/write transaction
semantics that we have omitted should be easily classifiable based on
the available primitives and HAT-incompatible anomaly prevention we
have already discussed.

@@

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared
  graphically. Directed edges represent ordering by model
  strength. Models that do not share a common ancestor can be
  simultaneously achieved, and the resulting availability is that of
  the weakest available model in the combination.}
\label{fig:hatcompared}
\end{figure}\vspace{-1em}

@@

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)
\end{align*}
If we want to guarantee both cut isolation and transactional atomicity
and the system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs
to read $a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute. Using client-side
caching can alleviate some of these challenges~\cite{bolton, swift},
but then the system becomes sticky high available.

>>>>

HAT systems provide transactions with transactional availability and
either high availability or sticky high availability. They offer
latency and availability benefits over traditional distributed
databases, yet they cannot achieve all the traditional semantics. In
this section, we delineate ACID, distributed consistency, and session
consistency levels which can be achieved with high availability
(transactional atomicity, variants of Repeatable Read isolation, and
many session guarantees), those with sticky high availability (read
your writes, PRAM and causal consistency) and the properties which
cannot be provided in a HAT system (those preventing Lost Update and
Write Skew, or with recency).  We present a full summary of these
results in Section~\ref{sec:hat-summary}.

@@

To begin, we present well-known semantics that can be achieved in HAT
systems. We offer proof-of-concept highly available algorithms, but
our goal is \textit{not} to provide optimal or even efficient
implementations. In our examples, we exclusively consider read and
write operations. We denote a write of value $v$ to data item $d$ as
$w_d(v)$ and a read from data item $d$ returning $v$ as $r_d(v)$ and
assume that all data items have the null value, $\bot$, at database
initialization. Unless otherwise specified, all transactions in the
examples commit.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
$PL-1$, requiring that each transaction's writes are ordered
consistently according to a single total order on transactions. This
prohibits Adya's phenomenon $G0$, also called ``Dirty
Writes''~\cite{adya}. We form a graph of transactions, with edges from
one transaction to another if the former writes to an object before
the latter writes to the same object. PL-1 prohibits cycles in this
graph, requiring some total order on writes. Accordingly, the final
database state cannot contain the ``earlier'' transaction's writes to
the data items: for example, in the below example, $T_3$ should
\emph{eventually} only read $a=b=1$ or $a=b=2$ but not $a=2, b=1$ or
$a=1, b=2$:
\begin{align*}
\small\vspace{-.5em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-.5em}
\end{align*}
Later properties will strengthen Read Uncommitted.

@@

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs. Implementations differ,
with some based on long-duration exclusive locks and short-duration
read locks~\cite{gray-isolation} and others based on multiple
versions. The implementations often provide recency and monotonicity
properties beyond the simple meaning from the name, which is what is
expressed in the implementation-agnostic definition: under Read
Committed, transactions should do not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's PL-2, and formalized by prohibiting Adya's $G1a-c$
(or ANSI's $P1$, or ``broad'' $P1$ (2.2) from Berenson et al.). For
instance, in the example below, $T_3$ should never see $a=1$, and, if
$T_2$ aborts, $T_3$ should never read $a=3$:
\begin{align*}
\small\vspace{-.5em}
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\vspace{-.5em}
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
transaction never writes uncommitted data to the database, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. This implementation does not provide recency or
monotonicity guarantees but satisfies the implementation-agnostic
definition.

``Repeatable Read'' isolation is a confusing property. Some IBM
products use the term for fully serializable
isolation~\cite{hat-hotos}. Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all interpret
Repeatable Read as providing serializability for all operations except
for predicate-based reads (and so Repeatable Read is identical to
serializability in a key-value store without predicate-based
access). In this model, the \textit{Phantom Problem}, whereby two
successive predicate-based reads return different
data~\cite{gray-isolation}, is the only kind of non-serializable
behavior allowed. In this paper, we reserve the use of ``Repeatable
Read'' to refer to these definitions, which we will soon show are not
HAT-compliant.

@@

A stronger achievable property requires consistent cuts to span both a
transaction's predicate reads (e.g., \texttt{SELECT WHERE}) as well as
its item reads.  We call this \textit{Predicate Cut
  Isolation} and it
prohibits Phantoms. Predicate Cut Isolation is also achievable in HAT
systems via similar multi-versioning as above.

@@

Perplexingly, discussions of TA are absent from existing ones of
weak isolation. This is perhaps again due to the single-node
context in which prior work was developed: on a single server (or a
fully replicated database), TA is achievable via lightweight locking
and/or local concurrency control over data items~\cite{gstore}. In
contrast, in a distributed environment, TA over arbitrary groups of
non-colocated items is considerably more difficult to achieve with
high availability: servers need to know when all of a transaction's
updates are present on their respective replicas. However, replicas do
not need to agree on when to reveal a new value to clients, which
would require consensus; accordingly, TA only requires reliable
broadcast (with some additional client metadata) and is
achievable in HAT systems.

To achieve TA, clients buffer their writes until commit, sending the
final write to each data item to each respective server. Along with
each write, clients attach a the transaction ID and a list of all data
items written to in the transaction, $L_v$. Upon receiving a new write
to item $d$, a server places the write into a \textit{pending} queue
and notifies all other servers responsible for a data item in $L_v$
that it has received a write to $d$ with the respective transaction
ID. Once a server receives acknowledgments with appropriate
transaction ID for all items in $L_v$, it moves the write from
\textit{pending} to the \textit{stable} database (which applies the
writes according to highest-ID-wins). Accordingly, all writes that
should be made atomic with a write in \textit{stable} are guaranteed
to be present on their respective servers, either in \textit{pending}
or \textit{stable}. In the presence of internal aborts, writes may
require 2 RTTs: one to check constraints and one to execute the above.

For reads, each client tracks which values it should read. It
maintains a mapping (vector) from data items to transaction IDs. Upon
transaction start, the client clears its mapping, and, when it
performs a read from a data item, it sends the server its current
value for the item. If the value is empty, the server reads from
\texttt{stable}. If the value is non-empty, the server either returns
the matching value from \texttt{pending} or a higher-ID value from
\texttt{pending}---along with $L_v$ for the write. When the client
receives a response, it updates its local vector with
$L_v$. Accordingly, even if transaction's writes to two keys $x$ and
$y$ are present on the $x$ server's \textit{stable} and the $y$
server's \textit{pending}, a client's reads from $x$ then $y$ will
return a transactionally atomic pair of values because the $y$ server
will know to check pending.

This implementation is entirely masterless and both reads and writes
do not block for coordination. One key property is that we do not, in
general, dictate when writes become visible. Accordingly, it is highly
available. However, we can optimize write visibility delay by ensuring
that clients are sticky with disjoint groups of servers. Once all
transaction values are \textit{pending} within a group of servers,
values can be made \textit{stable} (as opposed to once they are
present on all servers). For example, in a multi-datacenter setting,
all clients and servers within a datacenter may form a sticky group.

@@

Our models have not yet considered interactions \textit{across}
transactions, other than that there exists some arbitrary ordering on
them (via Read Uncommitted). In the distributed systems literature,
many useful \textit{safety} guarantees span multiple,
non-transactional operations. In particular, \textit{session
  guarantees} are used to describe guarantees across transactions
within a given \textit{session}, ``an abstraction for the sequence
of...operations performed during the execution of an
application''~\cite{sessionguarantees}. Informally a session describes
a context that should persist between transactions: for example, on a
social networking site, all of a user's transactions submitted between
``log in'' and ``log out'' operations might form a session.

@@

\vspace{.5em}\noindent\textit{{Monotonic reads}} requires that, within
a session, subsequent reads to a given object ``never return any
previous values''; reads from each item progress according to a total
order. The ordering of reads should respect any externally observable
total ordering on transactions.

\vspace{.5em}\noindent\textit{{Monotonic writes}} requires that each
session's writes become visible in the order they were submitted by
the client. Any order on transactions should also be consistent with
any precedence that a global observer would see.

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions
in succession:
\begin{align*}
\small\vspace{-.5em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-.5em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client executes $T_2$ against the same (partitioned)
server, then it will be able to read its writes. However, if the
network topology shifts and the client can only contact a different
server that is partitioned from the server that executed $T_1$, then
the client will be unable to read its own writes and the system will
have to either stall indefinitely to allow the client to read her
writes (violating transactional availability) or will have to
sacrifice read your writes guarantees. However, if the client remains
sticky with the server that executed $T_1$, then we can disallow this
scenario. Accordingly, read your writes, and, by proxy, causal
consistency and PRAM require stickiness. Read your writes is provided
by default in a sticky system, while the remaining causality and PRAM
guarantees can be accomplished with the algorithms for achieving the
remaining session guarantees.

@@

\vspace{0.5em}
\noindent{\textbf{Consistency}} A HAT system can make limited
application-level consistency guarantees. It can often execute
commutative and logically monotonic~\cite{calm} operations without the
risk of invalidating application-level integrity constraints. We do
not attempt to sketch the entire space of \textit{application-level}
consistency properties that are achievable in HAT systems (see
Section~\ref{sec:futurework}) but we specifically evaluate TPC-C
transaction semantics under HAT consistency guarantees in
Section~\ref{sec:evaluation}.

\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
transactions' effects survive $F$ server faults requires at least
$F+1$ non-failing nodes.

@@

In the words of Berenson et al., \textit{Lost Update} occurs when one
transaction $T1$ reads a given data item, a second transaction $T2$
updates the same data item, then $T1$ modifies the data item based on
its original read of the data item, ``missing'' or ``losing'' $T2$'s
newer update. Consider a database containing only the following
transactions:
\begin{align*}
\small\vspace{-.5em}
T_1 &: r_x(a) w_x(a+2)
\\T_2 &: w_x(2)\vspace{-.5em}
\end{align*}
If $T_1$ reads $a=1$ but $T_2$'s write to $x$ precedes $T_1$'s write
operation, then the database will end up with $a=3$, a state that
could not have resulted in a serial execution due to from $T_2$'s
``Lost Update.''

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small\vspace{-.5em}
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)\vspace{-.5em}
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:evaluation}) but, as we show in the extended version of
  the paper~\cite{hat-tr}, the problem persists in the general case.}
To prevent this from happening, either $T_1$ or $T_2$ should not have
committed. Each client's respective server might try to detect that
another write occurred, but this requires knowing the version of the
latest write to $x$. In our example, this reduces to a requirement for
linearizability, which is, via Gilbert and Lynch's proof of the CAP
Theorem, provably unachievable with high
availability~\cite{gilbert-cap}.

@@

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation, and Cursor Stability guarantees are all unavailable.
Repeatable Read (defined by to Gray, Berenson et al., and Adya) and
One-Copy Serializability need to prevent both Lost Update and Write
Skew. These prevention requirements mean that these models are
inherently unavailable.

@@

Data storage systems make various recency guarantees on reads of data
items. As we have discussed, one of the most famous is
linearizability~\cite{herlihy-art}, which states that reads will
return the last completed write to a data item, and there are several
other (weaker) variants such as safe and regular register
semantics. When applied to transactional semantics, the combination of
one-copy serializability and linearizability is called \textit{strong
  (or strict) one-copy serializability}~\cite{adya} (e.g.,
Spanner~\cite{spanner}). It is also common, particularly in systems
that allow reading from masters and slaves, to provide a guarantee
such as ``read a version that is no more than five seconds out of
date'' or similar. Unfortunately, an indefinitely long partition can
force an available system to violate any recency bound, so recency
bounds are not enforceable by HAT systems.

@@

We summarize our results in Table~\ref{table:hatcompared}. A wide
range of isolation levels are achievable in HAT systems, including
transactional atomicity, cut isolation, and several session
guarantees. With sticky availability, a system can achieve read your
writes guarantees and PRAM and causal consistency. However, many other
prominent models, such as Snapshot Isolation, One-Copy
Serializability, and Strict Serializability cannot be achieved due to
the inability to prevent Lost Update and Write Skew phenomena.

@@

To the best of our knowledge, this is the first unification of
database ACID, distributed consistency, and session guarantee
models. Interestingly, strong one-copy serializability subsumes all
other models, while considering the (large) power set of all
compatible models (e.g., the diagram depicts 96 possible HAT
combinations) hints at the vast expanse of consistency models found in
the literature. This taxonomy is not exhaustive, but we believe it
lends substantial clarity into the relationships between a large
subset of the prominent ACID and distributed consistency
models. Additional read/write transaction semantics that we have
omitted should be easily classifiable based on the available
primitives and HAT-incompatible anomaly prevention we have already
discussed.

@@

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared
  graphically. Directed edges represent ordering by model
  strength. Models that do not share a common ancestor can be
  simultaneously achieved, and the resulting availability is that of
  the weakest available model in the combination.}
\label{fig:hatcompared}
\end{figure}

@@

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small\vspace{-.5em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-.5em}
\end{align*}
If we want to guarantee both cut isolation and transactional atomicity
and the system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs
to read $a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute. Using client-side
caching can alleviate some of these challenges~\cite{bolton, swift},
but then the system becomes sticky high available.

### 2013-04-23 writeup/hat-appendix.tex
editing pass

\begin{definition}[Predicate Cut Non-Isolation (N-ICI)]
A history $H$ exhibits phenomenon N-ICI if there exists a transaction
$T_i$ such that $P= \{p \mid T_i$ directly predicate-depends by $p$ on another
other transaction$\}$ and $\bigcap_{p\in P} P \neq \emptyset$.
\end{definition}

>>>>

\begin{definition}[Predicate Cut Non-Isolation (N-ICI)]
A history $H$ exhibits phenomenon N-ICI if there exists a transaction
$T_i$ such that $P= \{p \mid T_i$ directly predicate-read-depends by
$p$ on another other transaction$\}$ and $\bigcap_{p\in P} P \neq
\emptyset$. (TODO: this isn't necessarily it)
\end{definition}

### 2013-04-23 writeup/hat-vldb.tex
References purge
cleaning
tweaks
fixed
editing pass

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax

\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, draftwatermarkbottom, draftwatermarktop, amsthm}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{xspace}

@@

@@

\title{Highly Available Transactions: Virtues and Limitations}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

\begin{abstract}
To minimize network latency and remain online during partial failure
and network partitions, many modern distributed data storage systems
eschew transactional functionality, which provides strong semantic
guarantees for groups of multiple operations over multiple data
items. In this work, we consider the problem of providing Highly
Available Transactions (HATs), or transactional guarantees that do not
suffer from unavailability during system partitions or incur high
network latency.  We introduce a new taxonomy of highly available
systems and analyze existing ACID and distributed consistency
guarantees to identify which can and cannot be achieved in HAT
systems. We analytically and experimentally quantify the performance
and availability benefits of HAT systems as well as the associated
semantic compromises.
\end{abstract}

@@

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% vldb_sample.bib is the name of the Bibliography in this case
\footnotesize
\bibliography{hat-vldb}  
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
\pagebreak

\begin{appendix}
\input{availability-appendix.tex}

@@

>>>>

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}

\newcommand{\pbnote}[1]{{\color{red}{#1 ---P.B.}}\\}

\interfootnotelinepenalty=10000

\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, amsthm}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{xspace}

@@

\makeatletter
\def\th@plain{%
  \thm@notefont{}% same as heading font
  \itshape % body font
}
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother

@@

\title{Highly Available Transactions: Virtues and Limitations}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

\begin{abstract}
To minimize network latency and remain online during server failures
and network partitions, many modern distributed data storage systems
eschew transactional functionality, which provides strong semantic
guarantees for groups of multiple operations over multiple data
items. In this work, we consider the problem of providing Highly
Available Transactions (HATs), or transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. We
analytically and experimentally quantify the performance and
availability benefits of HATs as well as their necessary semantic
compromises.
\end{abstract}

@@

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% vldb_sample.bib is the name of the Bibliography in this case
%\footnotesize
\scriptsize
\balance

\bibliography{hat-vldb}

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
%\pagebreak
%
\pagebreak
\begin{appendix}
%\input{availability-appendix.tex}

@@

% megastore
% redblue
% gemini
% redblue
% granola
% bernstein-concurrency
% surviving-failures-bodik
% blooml
% google-availability
% cops
